Biden Projected to Win the 2024 US Presidential Elections

### 2024-03-16 paper/paper.qmd
discussion added

>>>>

# Discussion
Our analysis modeled the variables of interest from the survey data, followed by post-stratification focusing on key variables. We plotted the trends for the post-stratified voting preferences by age, state, and sex. These visualizations yielded valuable insights into voter behavior. Notably, our findings indicated an inclination within the population towards voting for Biden from the Democratic Party.

## Votes By States
In the graph, @fig-support-state, we notice that all states in the U.S. other than Rhode Island, Missouri, Oklahoma and Nevada are more inclined to vote for the Democratic Party. We see that as we go towards the Midwest and Southern Regions, the support for Biden increases as more than 50% of the state is likely to vote for them. This is similar to the findings in [@James_2016] that address the same prediction for 2024. Our model and the article predict that states that support Trump are from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.

Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like manufacturing and agriculture have defined the middle and Midwest regions of the United States. These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.

## Age and Sex
@fig-support-age showcases different age groups and what their voting preferences are. We noticed that the youngest group of voters (age 18-29) and mid-age adults (30-44) are more than 70% inclined to vote for Biden. The emphasis on social justice and LGBTQ+ rights among the younger generation may be the cause of this difference. The democratic party led by Biden prioritizes these issues and speaks for the justice of the minorities [@Brownstein_2023]. Since there are more young people who identify as sexual minorities and they make up 10% of the U.S. population [@Meyer_2024], there is a trend for them to support Biden and the Democratic Party.

The middle-aged generation i.e individuals of age group from 30-44 years old were inclined to voting for Biden as well. Middle-aged voters may align more closely with Biden's policy proposals on issues such as healthcare, economic recovery, climate change, and social justice [@Igielnik_2021] .

Additionally, we notice almost a 10% decline in the votes for Biden from older age groups (45+ years old). The National security and conservative principles, which are frequently connected to the Republican Party, may have an impact on the elder generation.

@fig-support-sex shows a slight bias towards the democratic party from the females than the male gender. Predictions suggest that women are 10% more likely to vote for the Democratic Party than men. Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally favoring progressive policies over those of males [@Igielnik_2021]. This discrepancy is sometimes linked to the fact that women prioritize some  issues more than men do, with women giving social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program [@Igielnik_2021]. With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign seems to have connected more deeply with female voters, resulting in higher levels of support.

## General conclusion
@fig-support-general shows that in general the population of the United States is a lot more inclined to vote for Biden. We see a drastic difference in opinion of almost 40% i.e. people are 40% more likely to vote for Biden than vote for Trump. This conclusion highlighted the final prediction from our analysis. Furthermore, it sheds light on the prevailing social values and the direction individuals envision for their society. Some of the key issues that Americans wanted to see change were the healthcare reforms, climate change actions and racial justice and social equity. They wanted better national response to the COVID-19 pandemic, more ambitious climate policies and investments in renewable energy and lastly, a demand for reforms addressing systemic racism and criminal justice [@Brownstein_2023].

## Contributions to Knowledge:

By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences [@Gelman].

The utilization of post-stratification techniques in our analysis allowed for a more granular examination of voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.

Lastly, it highlights the high support of the Democratic Party led by Biden from the younger generations, females and people of the midwest and southern states.
